Semi-supervised low-rank logistic regression for high-dimensional neuroimaging data

### 2015-06-02 paper_nips/paper.tex
sneaked in BTs last comments

\author{Danilo Bzdok, Michael Eickenberg, Olivier Grisel,
  Bertrand Thirion,
  Ga\"el Varoquaux \\\\\textbf{\textit{email:} }firstname.lastname@inria.fr}

@@

\begin{abstract}
Imaging neuroscience links human behavior to aspects of brain
biology in ever-increasing datasets.
%
Existing neuroimaging methods typically perform either discovery of unknown
neural structure or testing of neural structure associated with mental tasks.
%
However, testing hypotheses on the neural correlates of an increasing number of mental tasks makes it crucial to represent the observations adequately.
%
We therefore propose to optimize
representation and task classification in
a unified statistical learning problem.
%
A multinomial logistic regression is introduced that is
constrained by joint rank with a representation-enforcing
autoencoder.
%
We show that this approach yields more accurate and more interpretable
neural models of psychological tasks in a reference dataset, and better generalization to other datasets.
%

\textbf{keywords}: dimensionality reduction; semi-supervised learning;
fMRI; systems neuroscience

@@

Both these biological properties can be conjointly exploited in an
approach that is
mixed (i.e., using task and rest data)
and 
low-rank (i.e., performing network decomposition).
%
The integration of brain-network discovery into
supervised classification can yield a semi-supervised learning
framework. The most relevant
neurobiological structure should hence be identified
for the prediction problem at hand.
%
Autoencoders suggest themselves because they can emulate
variants of most unsupervised learning algorithms,
including PCA, SPCA, and ICA \cite{hinton06}.
\begin{wrapfigure}{r}{0.40\textwidth}
  \centering
  %\captionsetup{labelformat=nonumber}
    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
  \vspace{-0.7cm}
  \caption {\textbf{Architecture.}
  Linear autoencoders find an optimal compression of 79,941 brain voxels
  into $n$ unknown activity patterns by improving reconstruction from them.
  The decomposition matrix equates with the bottleneck of
  a factored logistic regression.
  Supervised multi-class learning on task data ($X_{task}$)
  is thus guided by
  unsupervised decomposition of rest data ($X_{rest}$).
  }
\end{wrapfigure}
%
Autoencoders
are often one-layered learning models 
that condense the input data to
local and global 
representations
by optimizing reconstruction from them.
%
% XXX Michael: In fact, they behave like a non-centered PCA, 
% XXX which amounts to just doing an SVD
% XXX which btw is NOT a convex optimization problem, although easy to 
% XXX calculate and with a global optimum.
% XXX Autoencoders are also not one-layered by default. Let's make sure that
% XXX we don't trip up by getting the technicalities wrong and getting 
% XXX reviewers annoyed at us because of it.
% DanilO: cool thanks. That is the stuff, we need.
They behave like a (truncated, non-centered) PCA
in case of one linear hidden layer and a squared error loss
\cite{baldi1989neural}.
This architecture yields an optimization objective
with unique global minimum.
Autoencoders behave like a SPCA if shrinkage terms are added to the
model weights in the optimization objective.
Moreover, they have the characteristics of an ICA in case of tied weights
and adding a nonlinear convex function
at the first layer \cite{le2011ica}.
% XXX What convex transformation? Why is that relevant? Convexity is only
% XXX relevant for the loss
% Danilo: it's about adding a neural-network-like nonconvex function for
% first-layer activation
These authors further demonstrated that ICA, sparse autoencoders, and 
sparse coding are mathematically equivalent
under mild conditions.
%
Hence, autoencoders can flexibly project the neuroimaging data
onto the main axes of variation and thus % XXX Michael <- that's PCA
reverse-engineer properties of the data-generating
processes in the brain \cite{olshausen96}.
% XXX Michael ^ I don't get that last part nor the citation.
\textbf{XXX Bertrand: neither do I. I think you exxagerate.}
% Danilo: that is the magic

In the present investigation,
a linear autoencoder will be fit to
(unlabeled) rest data and
integrated as a rank-reducing bottleneck
into a multinomial
logistic regression fit to
(labeled) task data.
We can then solve the unsupervised data representation and the
supervised classification in an identical statistical learning
problem.
%
From the perspective of dictionary learning, the first layer represents
% XXX Michael: This is the crucial difference: The first layer is an analysis
% XXX operation: it extracts information in some way. It may be extracting
% XXX activations for basis functions, in which case it would be the pseudo-
% XXX inverse of those basis functions. But unless they are orthogonal, these
% XXX weights definitely don't represent basis functions themselves. It is 
% XXX the second layer of the autoencoder that represents basis functions. If
% XXX weights are tied, they are one and the same thing, with interesting
% XXX implications. But in general, only the second layer represents a basis.
projectors to the discovered set of basis functions
whose linear combinations are learned
by the second layer \cite{olshausen96}.
% XXX Michael: ^ This needs to be changed around: This first layer obtains
% XXX activations, the second layer weights are basis functions which are
% XXX multiplied to these activations
%
Neurobiologically, this allows 
delineating a low-dimensional manifold
of brain network patterns and then 
distinguishing mental tasks
by their most discriminative linear combinations.
%
Theoretically, a reduction in model variance should be achieved by
resting-state autoencoders that
favor the most neurobiologically
valid models in the hypothesis set.
%
In summary,
the important modes of variation in brain dynamics and
the neural correlates subserving mental operations
have mostly been studied in isolation.
We provide a principled computational framework to link these previously
unconnected domains of systems neuroscience.

\textbf{XXX Bertrand: I think that you could emphasize the problem that we are always data-poor, so that the set of representations that  can be extracted from SPMs taken from  a few tens of subjects is limited. What you propose it to extend it by i) co-analysing many problems simultaneously (multi-task) and using unlabled data that spana space of meaningful configurations
}
\textbf{XXX Bertrand: We need a sentence like: Our contribution is 1. XXX 2.YYY etc.}

@@

XXX (Bertrand): You can simplify data description if needed for the sake of place. It's OK as long as you can refer to a more complete publication describing the data.

@@

\paragraph{Optimization.}
The common objective was optimized
in the neuroimaging data
by gradient descent in the parameters
of the 
semi-supervised low-rank logistic regression (SS-LR LogReg).
The required gradients are easily obtained by using the chain rule to
backpropagate error derivatives through the linear network.
As solver, we chose \textit{rmsprop} \cite{rmsprop}.
%, a mini-batch version of rprop.
This procedure dictates an adaptive learning rate
for each model parameter by
scaled gradients from a running average.
Gradient normalization by \textit{rmsprop}
is known to effectively exploit curvature information.
We opted for a small batch size of $100$, given the high degree of
redundancy in $X_{rest}$ and $X_{task}$.
The matrix parameters were initalized by Gaussian random values multiplied
by a gain of $0.004$. 
%

XXX: Sorry, I don't like magic numbers. Please explain where this comes from
 ? Maybe $\sqrt{1 / p}$ ?

%
The bias parameters were initalized to zero.
With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
The normalization factor and the update rule for $\theta$
are given by
% XXX Michael: It is crucial, but nowhere mentioned, that the optimization is
% XXX done in an alternating manner. The whole structure relies on this, by
% XXX construction it is vulnerable to optimizing too much towards labels or
% XXX rest at a time. I also identify this as one of the major weak points
% XXX of the method, so it needs to be stated in a way that is not easily
% XXX attackable.
% XXX Maybe we could also talk less about RMSprop.
%
% XXX Danilo: I think we are not alternating, Xrest and Xtask go in
% simultaneously at every bach in every epoch

\begin{eqnarray}
  \begin{split}
    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
\\
\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
  \end{split}
\end{eqnarray}
% XXX Michael: Why, in the formula, do we have two different notations for
% XXX partial derivatives? \partial and \nabla

% XXX Michael: Probably remove the explanations following and just state values
% XXX: Bertrand: what is f ??
where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
$0.9$ to deemphasize the magnitude of the gradient.
Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
The hyper-parameter $\alpha$ was set to $0.00001$ by prior studies and $\epsilon$ was set to $10^{-6}$.
%
Note that we have also experimented with other solvers
(stochastic gradient descent, adadelta, and adagrad) but found that
\textit{rmsprop} converged faster and with higher generalization performance.

@@

\paragraph{Implementation.}
The analyses were performed in Python.
We used \textit{nilearn} to handle
the high-dimensional neuroimaging data 
\cite{abrah14}
and
\textit{Theano} for automatic, numerically stable
differentiation of symbolic computation graphs
\cite{bastien2012theano, bergstra2010theano}.
All Python scripts that generated the results are
accessible online for reproducibility and reuse
\url{http://github.com/anonymous/anonymous}.
% url{http://github.com/banilo/nips2015}.

@@

\begin{table}[h]
 \centering
 \resizebox{0.5\textwidth}{!}{%
 \begin{tabular}{lccc}
 \hline
 \textit{n}               & ICA + LogReg & SPCA + LogReg & SS-LR LogReg \\ \hline
 \multicolumn{1}{l|}{5}   & 37,53 \%     & 32,19 \%      & 96,50 \%     \\
 \multicolumn{1}{l|}{20}  & 80,98 \%     & 78,15 \%      & 97,33 \%     \\
 \multicolumn{1}{l|}{50}  & 84,19 \%     & 83,97 \%      & 97,69 \%     \\
 \multicolumn{1}{l|}{100} & 87,28 \%     & 82,19 \%      & 97,80 \%     \\ \hline
 \end{tabular}
 }
 \vspace{-0.2cm}
 \caption{Serial versus parallel dimensionality reduction and classification
XXX: only one decimal is meaningful}
 \label{table_one}
 \end{table}

\paragraph{Model performance.}
SS-LR LogReg was subsequently trained ($500$ epochs) across parameter choices
for the hidden components ($n=5, 20, 100$) and
the balance between autoencoder and logistic regression
($\lambda=0.25, 0.5, 0.75, 1$).
Assuming 5 latent directions of variation should yield models with
higher bias and smaller variance then SS-LR LogReg with 100 latent directions.
%
Given the 18-class problem of HCP, setting $\lambda$ to $0$
consistently yields generalization performance
at chance-level (5,6\%) because
only the unsupervised layer of the estimator is optimized.
%
At each epoch (i.e., iteration over the data),
the out-of-sample performance of the trained classifier
was assessed on 20\% of unseen HCP data.
Additionally, the ``out-of-dataset performance'' of the learned decomposition
was assessed by using it as dimensionality reduction of an
independent labeled dataset (i.e., ARCHI) and conducting ordinary
logistic regression on the ensuing component loadings.

\begin{table}[h]
  \centering
  \resizebox{1.00\textwidth}{!}{%
\begin{tabular}{l|cccc|cccc|cccc|cccc|}
  \hline
                            & \multicolumn{4}{c|}{\textit{n = 5}}                                                                                                            & \multicolumn{4}{c|}{\textit{n = 20}}                                                                                                                   & \multicolumn{4}{c}{\textit{n = 100}}                                                       \\ \hline
  \textit{}                 & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
  Out-of-sample\\accuracy   & \textit{88,90\%}                   & 95,13\%                            & \textbf{96,49\%}                  & 95,72\%                          & 97,44\%                              & \textbf{97,80\%}                    & 97,33\%                              & \textit{97,32\%}                   & 97,21\%          & \textit{97,03\%} & \textbf{97,80\%} & 97,38\%                            \\
  Precision (mean)          & \textit{86,98\%}                   & 94,86\%                            & \textbf{96,28\%}                  & 95,36\%                          & \textbf{97,38\%}                     & 97,06\%                             & \textit{96,98\%}                     & 97,02\%                            & 96,90\%          & \textit{96,47\%} & \textbf{97,50\%} & 96,91\%                            \\
  Recall (mean)             & \textit{88,27\%}                   & 95,19\%                            & \textbf{96,57\%}                  & 95,70\%                          & \textbf{97,51\%}                     & 97,50\%                             & \textit{97,35\%}                     & 97,36\%                            & 97,20\%          & \textit{97,15\%} & \textbf{97,88\%} & 97,37\%                            \\
  F1 score (mean)           & \textit{86,64\%}                   & 94,89\%                            & \textbf{96,38\%}                  & 95,43\%                          & \textbf{97,42\%}                     & 97,22\%                             & \textit{97,11\%}                     & 97,13\%                            & 97,00\%          & \textit{96,71\%} & \textbf{97,65\%} & 97,07\%                            \\ \hline
  Out-of-dataset \\accuracy & 60,83\%                            & \textit{54,30\%}                   & 60,69\%                           & \textbf{62,92\%}                 & 79,72\%                              & \textbf{81,94\%} & 79,72\%          & \textit{79,44\%}                     & \textbf{82,08\%} & 81,66\%         & 81,25\%          & \textit{75,83\%}                   \\ \hline
  \end{tabular}
}
\vspace{-0.2cm}
 \caption{Performance of SS-LR LogReg across model parameter choices. XXX: recall what the chance-level performance is.XXX: please add lambda=0}
  \label{table_two}
\end{table}

We made several noteworthy observations (Table \ref{table_two}).
%
First, the purely supervised estimator ($\lambda=1$) achieved in no
instance the best accuracy, precision, recall or f1 scores on HCP data.
Classification by low-rank logistic regression is therefore facilitated by
imposing structure from the unlabeled rest data.
%
Second, the higher the number of latent components $n$,
the higher the out-of-dataset performance with small choices for $\lambda$.
This suggests that the presence of more rest-data-inspired hidden components
results in more effective feature representations in unrelated task data.
%
Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
decomposition matrix ($\lambda=0$) resulted in
noninferior out-of-dataset performance
of X and Y, respectively (not shown in Table \ref{table_two} XXX: please show it).
This confirms that guiding model learning by task-unrelated structure
extract features of general relevance
beyond the supervised problem at hand.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/accuracies.png}
\end{centering}
\vspace{-0.5cm}
\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
Depicts f1 scores for prediction of each of 38 classes across training epochs.
Ordinary logistic regression operating in voxel space (\textit{left plot})
converged faster but performed worse than
low-rank logistic regression for few
(\textit{middle plot}) and many (\textit{right plot}) latent modes.
Autoencoder or rest data were not used for these analyses
($\lambda=1$).
Hence, projecting the input data into a reduced space for classification
yields higher class separability.
XXX: ugly figure: use larger font for labels and titles.
XXX: includes pdf figures, not a png file.
}
\label{fig_dimred}
\end{figure}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/semisup_both.png}
\end{centering}
\vspace{-0.5cm}
\caption{\textbf{Effect of richness in rest data}
Gradient descent was performed on 2000 task and 2000 rest images.
These were drawn with replacement from identical versus varying
quantities of task versus rest maps, at the begining of each epoch.
}
\label{fig_semisup}
\end{figure}

@@

\paragraph{Acknowledgment}
{\small
Data were provided by the Human Connectome Project. The study was supported
by the German National Academic Foundation (D.B.).
XXX: Note that this will be removed at submission for the sake of anonymization.
}

>>>>

\author{Danilo Bzdok, Michael Eickenberg, Olivier Grisel,
  Bertrand Thirion,
  Ga\"el Varoquaux \\\textbf{\textit{email:} }firstname.lastname@inria.fr}

@@

\begin{abstract}
Imaging neuroscience links human behavior to aspects of brain
biology in ever-increasing datasets.
%
Existing neuroimaging methods typically perform either discovery of unknown
neural structure or testing of neural structure associated with mental tasks.
%
However, testing hypotheses on the neural correlates of an
increasing number of mental tasks makes it crucial to
represent the observations adequately.
%
We therefore propose to optimize
representation and task classification in
a unified statistical learning problem.
%
A multinomial logistic regression is introduced that is
constrained by joint rank with a representation-enforcing
autoencoder.
%
We show that this approach yields more accurate and interpretable
neural models of psychological tasks in a reference dataset,
as well as better generalization to other datasets.
%

% Danilo: NIPS does actually not seem to have keywords
%\textbf{\\keywords}: curse of dimensionality; semi-supervised learning;
%fMRI; systems neuroscience

@@

Both these biological properties can be conjointly exploited in an
approach that is
mixed (i.e., using task and rest data)
and 
low-rank (i.e., performing network decomposition).
%
The integration of brain-network discovery into
supervised classification can yield a semi-supervised learning
framework. The most relevant
neurobiological structure should hence be identified
for the prediction problem at hand.
%
Autoencoders suggest themselves because they can emulate
variants of most unsupervised learning algorithms,
including PCA, SPCA, and ICA \cite{hinton06}.
\begin{wrapfigure}{r}{0.40\textwidth}
  \centering
  %\captionsetup{labelformat=nonumber}
    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
  \vspace{-0.7cm}
  \caption {\textbf{Architecture.}
  Linear autoencoders find an optimal compression of 79,941 brain voxels
  into $n$ unknown activity patterns by improving reconstruction from them.
  The decomposition matrix equates with the bottleneck of
  a factored logistic regression.
  Supervised multi-class learning on task data ($X_{task}$)
  is thus guided by
  unsupervised decomposition of rest data ($X_{rest}$).
  }
\end{wrapfigure}
%
Autoencoders
are often one-layered learning models 
that condense the input data to
local and global 
representations
by optimizing reconstruction from them.
%
% XXX Michael: In fact, they behave like a non-centered PCA, 
% XXX which amounts to just doing an SVD
% XXX which btw is NOT a convex optimization problem, although easy to 
% XXX calculate and with a global optimum.
% XXX Autoencoders are also not one-layered by default. Let's make sure that
% XXX we don't trip up by getting the technicalities wrong and getting 
% XXX reviewers annoyed at us because of it.
% DanilO: cool thanks. That is the stuff, we need.
They behave like a (truncated, non-centered) PCA
in case of one linear hidden layer and a squared error loss
\cite{baldi1989neural}.
This architecture yields an optimization objective
with unique global minimum.
Autoencoders behave like a SPCA if shrinkage terms are added to the
model weights in the optimization objective.
Moreover, they have the characteristics of an ICA in case of tied weights
and adding a nonlinear convex function
at the first layer \cite{le2011ica}.
% XXX What convex transformation? Why is that relevant? Convexity is only
% XXX relevant for the loss
% Danilo: it's about adding a neural-network-like nonconvex function for
% first-layer activation
These authors further demonstrated that ICA, sparse autoencoders, and 
sparse coding are mathematically equivalent
under mild conditions.
%
Hence, autoencoders can flexibly project the neuroimaging data
onto the main directions of variation.

In the present investigation,
a linear autoencoder will be fit to
(unlabeled) rest data and
integrated as a rank-reducing bottleneck
into a multinomial
logistic regression fit to
(labeled) task data.
We can then solve the unsupervised data representation and the
supervised classification, previously studied in isolation,
in an identical statistical learning problem.
%
From the perspective of dictionary learning, the first layer represents
% XXX Michael: This is the crucial difference: The first layer is an analysis
% XXX operation: it extracts information in some way. It may be extracting
% XXX activations for basis functions, in which case it would be the pseudo-
% XXX inverse of those basis functions. But unless they are orthogonal, these
% XXX weights definitely don't represent basis functions themselves. It is 
% XXX the second layer of the autoencoder that represents basis functions. If
% XXX weights are tied, they are one and the same thing, with interesting
% XXX implications. But in general, only the second layer represents a basis.
projectors to the discovered set of basis functions
whose linear combinations are learned
by the second layer \cite{olshausen96}.
% XXX Michael: ^ This needs to be changed around: This first layer obtains
% XXX activations, the second layer weights are basis functions which are
% XXX multiplied to these activations
%
Neurobiologically, this allows 
delineating a low-dimensional manifold
of brain network patterns and then 
distinguishing mental tasks
by their most discriminative linear combinations.
%
Theoretically, a reduction in model variance should be achieved by
resting-state autoencoders that
favor the most neurobiologically
valid models in the hypothesis set.
%
Practically, neuroimaging research is frequently suffering from
data scarcity. This limits the set of representations that can be
extracted from GLM analyses based on few participants.
%
We therefore contribute a computational framework that
1.) co-analyzes many problems simultaneously
(thus find shared representations by multi-task learning)
and
2.) exploits unlabeled data
(that span a space of meaningful configurations).

@@

@@

\paragraph{Optimization.}
The common objective was optimized
in the neuroimaging data
by gradient descent in the parameters
of the 
semi-supervised low-rank logistic regression (SS-LR LogReg).
The required gradients are easily obtained by using the chain rule to
backpropagate error derivatives through the linear network.
As solver, we chose \textit{rmsprop} \cite{rmsprop}.
%, a mini-batch version of rprop.
This procedure dictates an adaptive learning rate
for each model parameter by
scaled gradients from a running average.
Gradient normalization by \textit{rmsprop}
is known to effectively exploit curvature information.
We opted for a small batch size of $100$, given the high degree of
redundancy in $X_{rest}$ and $X_{task}$.
The matrix parameters were initalized by Gaussian random values multiplied
by a gain of $0.004$. 
%
The bias parameters were initalized to zero.
With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
The normalization factor and the update rule for $\theta$
are given by
% XXX Michael: It is crucial, but nowhere mentioned, that the optimization is
% XXX done in an alternating manner. The whole structure relies on this, by
% XXX construction it is vulnerable to optimizing too much towards labels or
% XXX rest at a time. I also identify this as one of the major weak points
% XXX of the method, so it needs to be stated in a way that is not easily
% XXX attackable.
% XXX Maybe we could also talk less about RMSprop.
%
% XXX Danilo: I think we are not alternating, Xrest and Xtask go in
% simultaneously at every bach in every epoch

\begin{eqnarray}
  \begin{split}
    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial \mathcal{L}}{\partial \theta}\right)^2
%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
\\
\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
  \end{split}
\end{eqnarray}
% XXX Michael: Why, in the formula, do we have two different notations for
% XXX partial derivatives? \partial and \nabla

% XXX Michael: Probably remove the explanations following and just state values
\textbf{XXX: Bertrand: what is f ?? Danilo: replaced f by loss L}

where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
$0.9$ to deemphasize the magnitude of the gradient.
Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
The hyper-parameter $\alpha$ was set to $0.00001$ by prior studies and $\epsilon$ was set to $10^{-6}$.
%
Note that we have also experimented with other solvers
(stochastic gradient descent, adadelta, and adagrad) but found that
\textit{rmsprop} converged faster and with higher generalization performance.

@@

\paragraph{Implementation.}
The analyses were performed in Python.
We used \textit{nilearn} to handle
the large quantities of neuroimaging data 
\cite{abrah14}
and
\textit{Theano} for automatic, numerically stable
differentiation of symbolic computation graphs
\cite{bastien2012theano, bergstra2010theano}.
All Python scripts that generated the results are
accessible online for reproducibility and reuse
\url{http://github.com/anonymous/anonymous}.
% url{http://github.com/banilo/nips2015}.

@@

\begin{table}[h]
 \centering
 \resizebox{0.5\textwidth}{!}{%
 \begin{tabular}{lccc}
 \hline
 \textit{n}               & ICA + LogReg & SPCA + LogReg & SS-LR LogReg \\ \hline
 \multicolumn{1}{l|}{5}   & 37,5 \%     & 32,2 \%      & 96,5 \%     \\
 \multicolumn{1}{l|}{20}  & 81,0 \%     & 78,2 \%      & 97,3 \%     \\
 \multicolumn{1}{l|}{50}  & 84,2 \%     & 84,0 \%      & 97,7 \%     \\
 \multicolumn{1}{l|}{100} & 87,3 \%     & 82,2 \%      & 97,8 \%     \\ \hline
 \end{tabular}
 }
 \vspace{-0.2cm}
 \caption{\textbf{Serial versus parallel dimensionality reduction and classification.}
 Chance is at 5,6\%.
 }
 \label{table_one}
 \end{table}

\paragraph{Model performance.}
SS-LR LogReg was subsequently trained ($500$ epochs) across parameter choices
for the hidden components ($n=5, 20, 100$) and
the balance between autoencoder and logistic regression
($\lambda=0, 0.25, 0.5, 0.75, 1$).
Assuming 5 latent directions of variation should yield models with
higher bias and smaller variance then SS-LR LogReg with 100 latent directions.
%
Given the 18-class problem of HCP, setting $\lambda$ to $0$
consistently yields generalization performance
at chance-level (5,6\%) because
only the unsupervised layer of the estimator is optimized.
%
At each epoch (i.e., iteration over the data),
the out-of-sample performance of the trained classifier
was assessed on 20\% of unseen HCP data.
Additionally, the ``out-of-dataset performance'' of the learned decomposition
was assessed by using it as dimensionality reduction of an
independent labeled dataset (i.e., ARCHI) and conducting ordinary
logistic regression on the ensuing component loadings.

\begin{table}[h]
  \centering
  \resizebox{1.00\textwidth}{!}{%
\tabcolsep=0.05cm
\begin{tabular}{l|ccccc|ccccc|ccccc|}
  \hline
                            & \multicolumn{5}{c|}{\textit{n = 5}}                                                                                                                                                 & \multicolumn{5}{c|}{\textit{n = 20}}                                                                                                                                                      & \multicolumn{5}{c}{\textit{n = 100}}                                                                           \\ \hline
  \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
  Out-of-sample\\accuracy   & \textit{6,0\%}                     & 88,9\%                             & 95,1\%                            & \textbf{96,5\%}                    & 95,7\%                           & \textit{5,5\%}                   & 97,4\%                               & \textbf{97,8\%}                     & 97,3\%                               & 97,3\%                             & \textit{6,1\%}   & 97,2\%           & 97,0\%           & \textbf{97,8\%}  & 97,4\%                             \\
  Precision (mean)          & \textit{5,9\%}                     & 87,0\%                             & 94,9\%                            & \textbf{96,3\%}                    & 95,4\%                           & \textit{5,1\%}                   & \textbf{97,4\%}                      & 97,1\%                              & 97,0\%                               & 97,0\%                             & \textit{5,9\%}   & 96,9\%           & 96,5\%           & \textbf{97,5\%}  & 96,9\%                             \\
  Recall (mean)             & \textit{5,6\%}                     & 88,3\%                             & 95,2\%                            & \textbf{96,6\%}                    & 95,7\%                           & \textit{4,6\%}                   & \textbf{97,5\%}                      & 97,5\%                              & 97,4\%                               & 97,4\%                             & \textit{7,2\%}   & 97,2\%           & 97,2\%           & \textbf{97,9\%}  & 97,4\%                             \\
  F1 score (mean)           & \textit{4,1\%}                     & 86,6\%                             & 94,9\%                            & \textbf{96,4\%}                    & 95,4\%                           & \textit{3,8\%}                   & \textbf{97,4\%}                      & 97,2\%                              & 97,1\%                               & 97,1\%                             & \textit{5,3\%}   & 97,0\%           & 96,7\%           & \textbf{97,7\%}  & 97,2\%                             \\ \hline
  Out-of-study\\accuracy    & \textit{39,4\%}                    & 60,8\%                             & 54,3\%                            & 60,7\%                             & \textbf{62,9\%}                  & \textit{77,0\%}                  & 79,7\%                               & \textbf{81,9\%}                     & 79,7\%                               & 79,4\%                             & 79,2\%           & \textbf{82,2\%}  & 81,7\%           & 81,3\%           & \textit{75,8\%}                    \\ \hline
  \end{tabular}
}
\vspace{-0.2cm}
 \caption{\textbf{Performance of SS-LR LogReg across model parameter choices.}
 Chance is at 5,6\%.
 }
  \label{table_two}
\end{table}

We made several noteworthy observations (Table \ref{table_two}).
%
First, the purely supervised estimator ($\lambda=1$) achieved in no
instance the best accuracy, precision, recall or f1 scores on HCP data.
Classification by low-rank logistic regression is therefore facilitated by
imposing structure from the unlabeled rest data.
%
Second, the higher the number of latent components $n$,
the higher the out-of-study performance with small choices for $\lambda$.
This suggests that the presence of more rest-data-inspired hidden components
results in more effective feature representations in unrelated task data.
%
Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
decomposition matrix ($\lambda=0$) resulted in
noninferior out-of-study performance
of 77,0\% and 79,2\%,
respectively (Table \ref{table_two}).
This confirms that guiding model learning by task-unrelated structure
extracts features of general relevance
beyond the supervised problem at hand.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/accuracies.png}
\end{centering}
\vspace{-0.5cm}
\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
Depicts f1 scores for prediction of each of 38 classes across training epochs.
Ordinary logistic regression operating in voxel space (\textit{left plot})
converged faster but performed worse than
low-rank logistic regression for few
(\textit{middle plot}) and many (\textit{right plot}) latent modes.
Autoencoder or rest data were not used for these analyses
($\lambda=1$).
Hence, projecting the input data into a reduced space for classification
yields higher class separability.
Chance is at 2,6\%.
XXX: ugly figure: use larger font for labels and titles.
XXX: includes pdf figures, not a png file.
}
\label{fig_dimred}
\end{figure}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/semisup_both.png}
\end{centering}
\vspace{-0.5cm}
\caption{\textbf{Effect of richness in rest data}
Gradient descent was performed on 2000 task and 2000 rest images.
These were drawn with replacement from identical versus varying
quantities of task versus rest maps, at the begining of each epoch.
Chance is at 5,6\%.
}
\label{fig_semisup}
\end{figure}

@@

\paragraph{Acknowledgment}
{\small
Data were provided by the Human Connectome Project.
% The study was supported by the German National Academic Foundation (D.B.).
}
